### 2016-07-07 image_setup.tex
Image OS Files and Configuration Setup Section done

\label{sec:image_setup}
In this Section, we review the steps to create a common \ac{OS} image for
all the \ac{Raspi}s. The image setup is composed of three major steps:
Select and download the \ac{OS} image file, alter the image structure and
configure the \ac{OS} files. We proceed to detail all these steps providing
brief discussions to our setup choices when required. To perform these
steps, we indicate with command-line blocks the required sequential
commands to be typed by the testbed administrator in its \ac{PC} to obtain
the desired setting. In all the following sections, the number sign
(\#) and dollar sign (\$) will be used in the command blocks in the paper
to indicate if an operation needs to be run with root permissions or
common user permissions, respectively. The signs will be posted as the
terminal prompt in the command blocks.

@@

\subsection{Default OS Files and Configuration Scripts Setup}

In general, the \ac{Raspi}s should be setup as similar as possible. However,
some particularities exist to differentiate the devices in principle. Also,
scripts containing further configurations for the \ac{Raspi}s are desirable
to be distributed as part of the common image. Therefore, here we present
the steps to setup basic properties of the \ac{Raspi}s and distributing
configuration scripts to each of them through the image. First, we indicate
how to obtain and put our configuration scripts in the image. Later, we
describe the tasks performed by these configuration scripts. Nevertheless,
any testbed administrator might modify or include other tasks besides the
presented in this procedure according to his / her needs.

In our case, we have our configuration scripts stored in a \texttt{.zip}
file located in a web server. We first retrieve this compressed file with
\texttt{wget} and unzip it locally into our Raspbian lite image. These
commands and the output of the last one are shown as follows:

@@

Among the settings that we want to be different in the devices is
their hostname. The hostname helps the user to physically distinguish the
devices from each other. Thus, in our case we require the devices in our
testbed to have different hostnames. Moreover, besides this standard task,
other testbed administrators may need to also include other scripts to
configure their \ac{Raspi}s. Hence, we want to ensure that all the \ac{Raspi}
configuration scripts of any testbed administrators are obtained in a simple
way. Therefore, we ease this task by making the \ac{Raspi}s fetch the
configuration scripts during start up before running them.

Hence, we define the hostnames based on the \ac{MAC} addresses
of the \ac{Raspi}s wired Ethernet interface. Prior to this stage, the
\ac{MAC} address of a network card can be found using the command
\texttt{ifconfig} or \texttt{ip addr} on a given \ac{Raspi}.
We store the \ac{MAC} addresses and hostnames of the \ac{Raspi}s in the
configuration file \texttt{\${ROOTDIR}/home/pi/rasp\_config/nodes.csv}. A
sample of our file is shown as follows:

@@

The testbed administrator has to insert the \ac{MAC} addresses and hostnames
of his/her \ac{Raspi}s obtained previously in the format shown in the
configuration file. Once we have defined the hostnames in
\texttt{\${ROOTDIR}/home/pi/rasp\_config/nodes.csv}, we use the
\texttt{\${ROOTDIR}/home/pi/rasp\_config/set\_hostname}
\ac{Bash} script to have each \ac{Raspi} assign its hostname:

@@

The \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname} script (in
script lines): (1) tells the system to interpret the script using \ac{Bash},
(3-4) gets the path to the script itself and the list of hostnames, (5) gets
the \ac{MAC} address of the node itself, (6) gets the current hostname, (7)
gets the new hostname from the hostname list and (10-14) assigns the new
hostname to the \ac{Raspi} where the script will be executed.

It is possible that the testbed administrator requires to add other
scripts to configure the \ac{Raspi}s besides the ones introduced here.
This would be tedious to distribute to all \ac{Raspi}s at a later stage.
We ease this task by making the \ac{Raspi}s fetch all the configuration
scripts during start up from a \texttt{.zip} file before running the
configuration scripts. The below script attempts to download/update the
scripts:

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/update\_rasp\_config}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
#!/usr/bin/env bash

@@

\begin{itemize}
    \item Line 3-4: Is the url and zip file that should be downloaded
    \item Line 7-10: Downloads the new configuration files to \texttt{/tmp}
        and prints a warning in case of errors
    \item Line 13: Unzips the files to \texttt{/home/pi/} while ignoring existing files/directories
\end{itemize}

For the above scripts to work in the \ac{Raspi}s, it is required that the
\ac{Raspi} \ac{MAC} address is found in \texttt{nodes.csv}. Thus,
\texttt{rasp\_config.csv} needs to be unzipped, altered, and zipped again in
the \ac{HTTP} server from which the \ac{Raspi} retrieves the files from.
The testbed administrator will therefore need to also edit the \ac{URL} in the
script \texttt{update\_rasp\_config}.
%"nodes.csv" can be changed using a text editor. For vim, this could be done
%the following way:
%\begin{lstlisting}[]
%$ vim ${ROOTDIR}/home/pi/rasp_config/nodes.csv
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}

Finally, to actually make the \ac{Raspi}s change hostname, we have to make
each \ac{Raspi} call the above script when it starts up. Optionally,
also run the update script before running other scripts.
Do this with root permissions by inserting the line(s) below before
\texttt{"exit 0"} using your favorite editor:
%"\$\{ROOTDIR\}/etc/rc.local" using
%your favorite text editor..
%The file
%can be opened with your favorite text editor.

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
<@\textcolor{gray}{
-----------------------------------------------------}
\Reactivatenumber @>
...
bash /home/pi/rasp_config/update_rasp_config
bash /home/pi/rasp_config/set_hostname
...
exit 0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Configuring \texttt{rc.local} makes it really cumbersome for the testbed
administrator to add new scripts since it first needs to be included
in \texttt{rasp\_config.zip} and then each \ac{Raspi} needs to have an additional
line added to \texttt{rc.local} in the root image or in all memory cards depending on
the final testbed setup.

A solution to this is to make yet another script in \texttt{rasp\_config.zip}
that calls all the other scripts in the desired order. This would look like
below:
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/main}@>
<@\textcolor{gray}{
-----------------------------------------------------}
\Reactivatenumber @>
#!/usr/bin/env bash

bash /home/pi/rasp_config/set_hostname
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Now, to use the above script, \texttt{rc.local} needs to look like this:
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
<@\textcolor{gray}{
-----------------------------------------------------}
\Reactivatenumber @>
...
bash /home/pi/rasp_config/update_rasp_config
bash /home/pi/rasp_config/main
...
exit 0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
Notice that \texttt{set\_hostname} is now called by the \texttt{main}
script instead. The update script is still called directly. This ensures that
all configuration scripts are updated before executed. Changes to the update
script itself will first take effect at next system startup.
%Had the update script
%been called in \texttt{main} instead would mean that changes to the
%\texttt{main} script would also first have been available at next startup.

%This can be done by inserting a call to the script
%set\_hostname in rc.local before it exits.
%
%Call set\_hostname script at startup. We insert a line of code to call script in rc.local just after the initial comments (i.e. lines starting with \#).
%\begin{lstlisting}[]
%# line_number=$(egrep -n -m1 "(^[^#])|(^$)" ${ROOTDIR}/etc/rc.local | cut -d: -f1)
%# sed -i "${line_number}c bash /home/pi/rasp_config/set_hostname" ${ROOTDIR}/etc/rc.local
%\end{lstlisting}
%\FloatBarrier
%$ line="bash /home/pi/.rasp_config/set_hostname"
%# awk -v text="$line" '!/^#/ && !p {print text; p=1} 1' ${ROOTDIR}/etc/rc.local > <@ \Suppressnumber @>
%    ${ROOTDIR}/etc/rc.local <@ \Reactivatenumber @>

%WE NEED TO ADD A FEW COMMANDS TELLING HOW TO ZIP, MODIFY, AND UNZIP
%FOR THE READER TO BE ABLE TO UPLOAD A CONSTRUCT HIS/HER OWN BUNDLE.

%All other files could be updated as well.
It is not only possible to add and configure files to/within the image. One can
also install and update the packages within the image which will be illustrated
in the next section.

\subsection{Installing and updating the image}

>>>>

\label{sec:image_setup}
In this Section, we review the steps to create a common \ac{OS} image for
all the \ac{Raspi}s. The image setup is composed of three major steps:
Select and download the \ac{OS} image file, alter the image structure and
configure the \ac{OS} files. We proceed to detail all these steps providing
brief discussions to our setup choices when required. To perform these
steps, we indicate with command-line blocks the required sequential
commands to be typed by the testbed administrator in his / her \ac{PC}
to obtain the desired setting. In all the following sections, the number
sign (\#) and dollar sign (\$) will be used in the command blocks in the
paper to indicate if an operation needs to be run with root permissions or
common user permissions, respectively. The signs will be posted as the
terminal prompt in the command blocks.

@@

\subsection{Image OS Files and Configuration Scripts Setup}
In general, the \ac{Raspi}s should be setup as similar as possible. However,
some particularities exist to differentiate the devices in principle. Also,
scripts containing further configurations for the \ac{Raspi}s are desirable
to be distributed as part of the common image. Therefore, here we present
the steps to setup basic properties of the \ac{Raspi}s and distributing
configuration scripts to each of them through the image. For this, we first
indicate how to obtain and put our configuration scripts in the image. Later,
we describe the tasks performed by these configuration scripts. Finally, we
indicate how and in which order are the scripts executed to configure all the
devices. Nevertheless, any testbed administrator might modify or include other
tasks besides the presented in this procedure according to his / her needs
as we will show.

\subsubsection{Image Default Configuration Scripts Download}
\label{sec:configuration_files_download}
In our case, we have our default configuration scripts stored in a
\texttt{rasp\_config.zip} file located in the same \ac{URL} where the image
was retrieved from, i.e. the one in the environment variable \texttt{URL}.
We first manually download this compressed file with \texttt{wget} and
extract it locally into our Raspbian lite image. These commands and the
output of the last one are shown as follows:

@@

\subsubsection{Device Hostnames}
Among the settings that we want to be different in the devices is
their hostname. The hostname helps the user to physically distinguish the
devices from each other. Thus, in our case we require the devices in our
testbed to have different hostnames. Hence, we define the hostnames based
on the \ac{MAC} addresses of the \ac{Raspi}s wired Ethernet interface.

Prior to this stage, the \ac{MAC} address of a network card can be found
using the command \texttt{ifconfig} or \texttt{ip addr} on a given
\ac{Raspi}. We store the \ac{MAC} addresses and hostnames of the
\ac{Raspi}s in the configuration file
\texttt{\${ROOTDIR}/home/pi/rasp\_config/nodes.csv}. A sample of our file
is shown as follows:

@@

The testbed administrator has to insert the \ac{MAC} addresses and hostnames
of his/her \ac{Raspi}s obtained previously in the format shown in the
configuration file. In the file format, on each line there is a \ac{MAC}
address and the corresponding hostname for the given \ac{Raspi}. For
simplicity in the example above, we just include three \ac{Raspi}s.
This list file will be employed by the
\texttt{\${ROOTDIR}/home/pi/rasp\_config/set\_hostname}
\ac{Bash} script to assign the hostname of each \ac{Raspi}.
The script content is the following:

@@

The script (in lines): (1) tells the system to interpret the script
using \ac{Bash}, (3-4) gets the path to the script itself and the list of
hostnames, (5) gets the \ac{MAC} address of the node itself, (6) gets the
current hostname, (7) gets the new hostname from the hostname list and
(10-14) assigns the new hostname to the \ac{Raspi} where the script
will be executed.

\subsubsection{Updating Default Configuration Files and Scripts}
Besides the single script with its configuration file introduced up
to this point of our procedure, it is possible that the testbed
administrator may require to add other scripts to configure
his / her \ac{Raspi}s. Hence, we want
to ensure that all the \ac{Raspi} configuration scripts of any testbed
administrators are obtained in a simple way. We automate this task by
including the \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/update\_rasp\_config}
script in our procedure. The purpose of this script is to make all the
\ac{Raspi}s fetch all the configuration scripts located with the image
during a testbed start up.

In our case as the testbed administrator for presenting the procedure, we want
to fetch all our configuration scripts in
\texttt{\$\{URL\%/*\}/raspi\_config.zip}. Thus, the update script
\textit{automatically} downloads all the required configuration files in
\texttt{rasp\_config.zip} file from a remote location. This is the same that
we manually did earlier to get our files but this will be made in an
automated wat after booting up the system. This script contents are:

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/update\_rasp\_config}@>
<@\textcolor{gray}{
--------------------------------------------------------------------------------------------------}
\Reactivatenumber @>
#!/usr/bin/env bash

@@

The update script: (3-4) checks for the \ac{URL} and \texttt{.zip} file that
should be downloaded, (7-10) downloads the configuration files to \texttt{/tmp}
folder in the corresponding \ac{Raspi}. It also prints a warning in case of
errors and (13) unzips the files to \texttt{/home/pi/}, the \ac{Raspi} home
directory, while ignoring existing files and / or directories there.

For the above scripts to work in the \ac{Raspi}s, it is required that the
\ac{Raspi}s \ac{MAC} addresses are found in \texttt{nodes.csv}. Also, it
should be noted that for other testbed administrators besides ourselves,
the \ac{URL} for file fetching and the configuration
scripts themselves can be modified to fit their requirements. Thus, if
required for a testbed administrator, the \texttt{rasp\_config.zip}
will need to be edited to include all the required configuration files
and scripts. Also, it might be needed to edit the \ac{URL} in the script
\texttt{update\_rasp\_config} to store and fetch from a different location.
Nevertheless, both the \ac{URL} and configuration files presented here can be
used as a starting boilerplate if desired.

\subsubsection{Configuration Scripts Execution Order}
To actually make the \ac{Raspi}s change hostnames and any other considered
configurations, we have to make each \ac{Raspi} call the above scripts when
it starts up. After finishing the setup process, all the unzipped files
presented in Section~\ref{sec:configuration_files_download} should be
locally available at each \ac{Raspi} after getting the root filesystem.
Therefore, we first need to run the update script before running
any other configuration scripts. To do this after boot up, we include
a call for the update script in the image \texttt{\$\{ROOTDIR\}/etc/rc.local}
file before \texttt{exit 0} using a file editor. To do this, it is required
to edit this file as root. The resulting file should look like the following:

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
<@\textcolor{gray}{
---------------------------------------------}
\Reactivatenumber @>
...
bash /home/pi/rasp_config/update_rasp_config
...
exit 0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

If it is required to have more configuration scripts, adding them
in the \texttt{rc.local} file difficults its maintenance by the testbed
administrator since this needs to be both in the image and the
downloaded \texttt{rasp\_config.zip}. To avoid this problem, we include
the \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/main} script which
calls all other configuration scripts (besides \texttt{update\_rasp\_config})
in sequential order. This script content is the following:

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/main}@>
<@\textcolor{gray}{
---------------------------------------------------------------------}
\Reactivatenumber @>
#!/usr/bin/env bash

bash /home/pi/rasp_config/set_hostname
# Any other required configuration scripts...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

In this way, the automation process is simplify since we only need to modify
the image \texttt{\$\{ROOTDIR\}/etc/rc.local} file once to execute this script:

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
<@\textcolor{gray}{
-----------------------------------------------------}
\Reactivatenumber @>
...
bash /home/pi/rasp_config/update_rasp_config
bash /home/pi/rasp_config/main
...
exit 0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
Notice that \texttt{set\_hostname} is now called by the \texttt{main}
script instead. The update script is still called directly. This ensures that
all configuration scripts are updated before executed. Changes to the update
script itself will first take effect at next system startup.

\subsection{Image Package Updating}
Besides adding and configuring files within the image, the testbed
administrator can also install and update the software packages within
the image which will be illustrated in this subsection.
